Bayesian Insights into Olive Oil Pricing Dynamics in Canada

### 2024-12-04 paper/paper.qmd
Add files via upload

From the statistics presented in the table (see @fig-staCurrentPrice), the mean price of olive oil is approximately `$15.52`, with a median slightly higher at `$15.99`.  This suggests a slightly skewed distribution, which is confirmed by the histogram (@fig-frequCurrentPrice).  The standard deviation of prices is around `$5.99`, indicating a moderate variation in the prices of olive oil across different locations or brands.  The prices range from a minimum of `$4.97` to a maximum of `$42.99`, highlighting significant differences in pricing strategies or product types across stores.

@@

This boxplot ( @fig-VendorCurrentPrice) compares the current_price of olive oil products across different vendors: Loblaws, Metro, and Walmart. The variable vendor reflects the retail chain selling the product, and differences in pricing can provide insights into vendor-specific strategies and market positioning.

@@

The primary aim of this study is to forecast the future price trends of olive oil by gathering representative data across various market segments, including historical pricing, supply chain dynamics, consumer demand, and broader economic factors. By utilizing statistical modeling techniques and machine learning methods, this research aims to increase the accuracy of price predictions, which can provide valuable insights to retailers, consumers, and policymakers. The analysis will also evaluate the factors that influence price fluctuations, including macroeconomic variables, supply chain disruptions, and seasonal price changes [@Smith_Johnson_2023]. This study seeks to refine existing price prediction models and offer more reliable forecasts in an increasingly volatile market.
Data Collection and Sampling Strategy
To ensure broad representativeness, the study will collect price data from multiple sources, including retail supermarkets, online platforms, and supply chain entities. The data will span the past year to incorporate both seasonal trends and short-term market fluctuations. These factors, combined with promotional offers and other relevant economic variables, will be considered in the analysis. Sampling for this study will also include supplier, wholesaler, and importer price data to capture a comprehensive view of the olive oil pricing ecosystem [@Anderson_Davis_2022].
The sampling approach will combine time-series data collection with economic factors, such as inflation and exchange rate fluctuations. Additionally, production cost variables will be factored into the pricing models to ensure the reliability and generalizability of the findings. The estimated sample size for this study is 3,000 data points, which will be sufficient to perform trend analysis and model training with adequate statistical power [@Lee_Chang_2021].

@@

To ensure the accuracy and integrity of the data, several validation techniques will be employed. The first step will be data cleaning to remove missing values and outliers that could distort the analysis. Cross-validation of pricing data from different retailers and suppliers will be conducted to ensure consistency. Furthermore, the temporal accuracy of timestamps and the legitimacy of price fluctuations will be verified to minimize any discrepancies that might influence the predictive modeling process [@Chen_Li_Zhang_2020]. Data quality measures will also include real-time verification checks and cross-referencing with publicly available pricing data to increase the robustness of the results.
Data Analysis and Modeling Approach
For the data analysis, a combination of traditional statistical modeling and machine learning techniques will be utilized. Time-series forecasting using ARIMA models will serve as the baseline for predicting long-term price trends and seasonal fluctuations. This will be complemented by multiple regression models to identify the key determinants of price changes, such as supply chain disruptions and changes in production costs [@Taylor_Brown_2019]. To further refine the analysis, regularization techniques like ridge regression and Lasso regression will be employed to avoid multicollinearity and overfitting issues that could affect the robustness of the predictions.
In cases of nonlinear relationships and high-dimensional data, machine learning techniques such as random forests and support vector machines (SVM) will be applied to capture more complex patterns in price variation. Cross-validation will be implemented to ensure that the models generalize well, and evaluation metrics such as Mean Squared Error (MSE) and Mean Absolute Error (MAE) will be used to assess the model's predictive accuracy. Aggregating results from multiple models will ensure that the final predictions are stable and reliable [@Hastie_Tibshirani_Friedman_2009].

@@

The total budget for this study will be allocated across several key areas of research and development. Data collection and cleaning will account for $30,000, while the statistical analysis and model development will be allocated $40,000. The budget for computational resources, including cloud storage and processing, will amount to $15,000. Additionally, $10,000 will be set aside for the preparation of research reports and dissemination of results. Market research and expert consultation will be allocated $5,000. This budget distribution ensures the effective use of resources to achieve the study's objectives [@Peters_Adams_2023].
Research Implementation and Timeline
The study will be carried out over a three-month period, with regular updates provided on the price prediction outcomes. The data sources for this project will primarily include public retail platforms, supply chain databases, and price data provided by suppliers and wholesalers. This ensures that the research is based on a broad and current dataset, increasing the relevance and timeliness of the findings. The predicted olive oil prices will be periodically updated to reflect any significant market changes, and the final results will be presented in both written reports and interactive visualizations for easy interpretation by stakeholders [@Baker_2021].

>>>>

From the statistics presented in the table (@fig-staCurrentPrice), the mean price of olive oil is approximately `$15.52`, with a median slightly higher at `$15.99`.  This suggests a slightly skewed distribution, which is confirmed by the histogram (@fig-frequCurrentPrice).  The standard deviation of prices is around `$5.99`, indicating a moderate variation in the prices of olive oil across different locations or brands.  The prices range from a minimum of `$4.97` to a maximum of `$42.99`, highlighting significant differences in pricing strategies or product types across stores.

@@

This boxplot (@fig-VendorCurrentPrice) compares the current_price of olive oil products across different vendors: Loblaws, Metro, and Walmart. The variable vendor reflects the retail chain selling the product, and differences in pricing can provide insights into vendor-specific strategies and market positioning.

@@

The primary aim of this study is to forecast the future price trends of olive oil by gathering representative data across various market segments, including historical pricing, supply chain dynamics, consumer demand, and broader economic factors. By utilizing statistical modeling techniques and machine learning methods, this research aims to increase the accuracy of price predictions, which can provide valuable insights to retailers, consumers, and policymakers. The analysis will also evaluate the factors that influence price fluctuations, including macroeconomic variables, supply chain disruptions, and seasonal price changes. This study seeks to refine existing price prediction models and offer more reliable forecasts in an increasingly volatile market.
Data Collection and Sampling Strategy
To ensure broad representativeness, the study will collect price data from multiple sources, including retail supermarkets, online platforms, and supply chain entities. The data will span the past year to incorporate both seasonal trends and short-term market fluctuations. These factors, combined with promotional offers and other relevant economic variables, will be considered in the analysis. Sampling for this study will also include supplier, wholesaler, and importer price data to capture a comprehensive view of the olive oil pricing ecosystem [@Anderson_Davis_2022].
The sampling approach will combine time-series data collection with economic factors, such as inflation and exchange rate fluctuations. Additionally, production cost variables will be factored into the pricing models to ensure the reliability and generalizability of the findings. The estimated sample size for this study is 3,000 data points, which will be sufficient to perform trend analysis and model training with adequate statistical power [@Lee_Chang_2021].

@@

To ensure the accuracy and integrity of the data, several validation techniques will be employed. The first step will be data cleaning to remove missing values and outliers that could distort the analysis. Cross-validation of pricing data from different retailers and suppliers will be conducted to ensure consistency. Furthermore, the temporal accuracy of timestamps and the legitimacy of price fluctuations will be verified to minimize any discrepancies that might influence the predictive modeling process [@Chen_Li_Zhang_2020]. Data quality measures will also include real-time verification checks and cross-referencing with publicly available pricing data to increase the robustness of the results.
Data Analysis and Modeling Approach
For the data analysis, a combination of traditional statistical modeling and machine learning techniques will be utilized. Time-series forecasting using ARIMA models will serve as the baseline for predicting long-term price trends and seasonal fluctuations. This will be complemented by multiple regression models to identify the key determinants of price changes, such as supply chain disruptions and changes in production costs. To further refine the analysis, regularization techniques like ridge regression and Lasso regression will be employed to avoid multicollinearity and overfitting issues that could affect the robustness of the predictions.
In cases of nonlinear relationships and high-dimensional data, machine learning techniques such as random forests and support vector machines (SVM) will be applied to capture more complex patterns in price variation. Cross-validation will be implemented to ensure that the models generalize well, and evaluation metrics such as Mean Squared Error (MSE) and Mean Absolute Error (MAE) will be used to assess the model's predictive accuracy. Aggregating results from multiple models will ensure that the final predictions are stable and reliable [@Hastie_Tibshirani_Friedman_2009].

@@

The total budget for this study will be allocated across several key areas of research and development. Data collection and cleaning will account for $30,000, while the statistical analysis and model development will be allocated $40,000. The budget for computational resources, including cloud storage and processing, will amount to $15,000. Additionally, $10,000 will be set aside for the preparation of research reports and dissemination of results. Market research and expert consultation will be allocated $5,000. This budget distribution ensures the effective use of resources to achieve the study's objectives.
Research Implementation and Timeline
The study will be carried out over a three-month period, with regular updates provided on the price prediction outcomes. The data sources for this project will primarily include public retail platforms, supply chain databases, and price data provided by suppliers and wholesalers. This ensures that the research is based on a broad and current dataset, increasing the relevance and timeliness of the findings. The predicted olive oil prices will be periodically updated to reflect any significant market changes, and the final results will be presented in both written reports and interactive visualizations for easy interpretation by stakeholders [@Baker_2021].
